### 2019-05-19 README.rmd
updating rclean

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
original_files <- dir()
```

<!-- # ijtiff  <img src="man/figures/logo.png" height="140" align="right"> -->

<!-- Code status -->
[![Build Status](https://travis-ci.org/ProvTools/Rclean.svg?branch=master)](https://travis-ci.org/ProvTools/Rclean) 
[![Build Status](https://travis-ci.org/ProvTools/Rclean.svg?branch=dev)](https://travis-ci.org/ProvTools/Rclean)

<!-- R status -->
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/Rclean)](https://cran.r-project.org/package=Rclean)
![RStudio CRAN downloads](http://cranlogs.r-pkg.org/badges/grand-total/Rclean)
![RStudio CRAN monthly downloads](http://cranlogs.r-pkg.org/badges/Rclean)
[![Rdocumentation](http://www.rdocumentation.org/badges/version/Rclean)](http://www.rdocumentation.org/packages/Rclean)

<!-- Dev status -->
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)

<!-- Package Review -->
[![](https://badges.ropensci.org/300_status.svg)](https://github.com/ropensci/onboarding/issues/300)
[![status](http://joss.theoj.org/papers/334d80d5508056dc6e7e17c6fd3ed5a6/status.svg)](http://joss.theoj.org/papers/334d80d5508056dc6e7e17c6fd3ed5a6)

<!-- Archiving -->
[![DOI](https://zenodo.org/badge/102645585.svg)](https://zenodo.org/badge/latestdoi/102645585)

Clean up your code
==================

- Have you ever written a long script in R that conducts lots of
  analyses and wished that someone would come along and make it all
  clearer to understand and use?
- Well you’re not alone. 
- A recent survey of over 1500 scientists reported a crisis of
  reproducibility with "selective reporting" being the most cited
  contributing factor and 80% saying code availability is playing a
  role.
- [Rclean](https://github.com/ProvTools/Rclean) was created to help
  scientists more *easily* write "cleaner" code by providing a simple,
  rigorous way to isolate the minimal code you need in order to
  produce a specific result (e.g. object, plot, output written to
  disk).
- [Rclean](https://github.com/ProvTools/Rclean) is built on data
  provenance, a formal representation of a computation, and uses graph
  analysis to determine the minimal path from inputs to results.
- The goal is to help facilitate code transparency, reproduciblity,
  reusability and ultimately help scientists spend more time on
  research and less time on software.

Install and Setup
=================

You can install
[Rclean](https://cran.r-project.org/web/packages/Rclean/) from *CRAN*:

```R
install.packages("Rclean")
```

You can install the most up to date (beta) version easily with
[devtools](https://github.com/hadley/devtools):

```R
install.packages("devtools")
devtools::install_github("ProvTools/Rclean", ref = "dev")
```

Once installed, per usual R practice, just load the *Rclean*:

```R
library(Rclean)
```

Usage
=====

*Rclean* usage is simple. Have a script with code you want to clean
saved to disk.

Once you have your script and workspace setup, you can use
[Rclean](https://github.com/ProvTools/Rclean) to get clean chunks of a
larger script that produce specific results you want. We'll use the
micro.R script, which can be found inside the package repo in the
*exec* directory. The following example assumes that your current
working directory is *exec*.

First, you'll need to record information about the script you would
like to parse. [Rclean](https://github.com/ProvTools/Rclean) uses data
provenance to verify what lines of code depend on each other inside of
the larger script. We can use the
[provR](https://github.com/ProvTools/provR) package to generate
provenance. The next bit of code runs our script and saves the
provenance to memory, which we then pass to the `options` function, so
that [Rclean](https://github.com/ProvTools/Rclean) has access to it:

```R
prov.capture("micro.R")
options(prov.json = prov.json())
```

Or, if you have provenance saved as a text file, you can load it in
like this:

```R
options(prov.json = readLines("prov_micro.json"))
```

Now that we have the provenance loaded, we can start
cleaning. [Rclean](https://github.com/ProvTools/Rclean) will give us a
list of possible values we can get code for:

```R
clean()

```

You can then pick and choose from among these results and get the
essential code to produce the output, like so:

```R
clean(x)

```

Notice that the 'clean' function doesn't require you to quote your
results, it interprets all inputs as names of results.

In many cases, it's handy just to take a look at the isolated code,
but if you can also save the code for later use or sharing.

```R
my.code <- clean(x)
write.code(my.code, file = "x.R")

```

If you would like to copy your code to the clipboard, you can do that
by not specifying a file path.

```R
write.code(my.code)

```

Happy cleaning!

*Contributing*: if you would like to contribute, please read [[CONTRIBUTING.md]].