### 2019-06-03 README.Rmd
adds tests for clean

---
output: md_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
original_files <- dir()
```

<!-- # ijtiff  <img src="man/figures/logo.png" height="140" align="right"> -->

<!-- Code status -->
[![Build Status](https://travis-ci.org/ProvTools/Rclean.svg?branch=master)](https://travis-ci.org/ProvTools/Rclean) 
[![Build Status](https://travis-ci.org/ProvTools/Rclean.svg?branch=dev)](https://travis-ci.org/ProvTools/Rclean)

<!-- R status -->
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/Rclean)](https://cran.r-project.org/package=Rclean)
![RStudio CRAN downloads](http://cranlogs.r-pkg.org/badges/grand-total/Rclean)
![RStudio CRAN monthly downloads](http://cranlogs.r-pkg.org/badges/Rclean)
[![Rdocumentation](http://www.rdocumentation.org/badges/version/Rclean)](http://www.rdocumentation.org/packages/Rclean)

<!-- Dev status -->
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)

<!-- Package Review -->
[![](https://badges.ropensci.org/300_status.svg)](https://github.com/ropensci/onboarding/issues/300)
[![status](http://joss.theoj.org/papers/334d80d5508056dc6e7e17c6fd3ed5a6/status.svg)](http://joss.theoj.org/papers/334d80d5508056dc6e7e17c6fd3ed5a6)

<!-- Archiving -->
[![DOI](https://zenodo.org/badge/102645585.svg)](https://zenodo.org/badge/latestdoi/102645585)

Clean up your code
==================

- Have you ever written a long script in R that conducts lots of
  analyses and wished that someone would come along and make it all
  clearer to understand and use?
- Well you’re not alone. 
- A recent survey of over 1500 scientists reported a crisis of
  reproducibility with "selective reporting" being the most cited
  contributing factor and 80% saying code availability is playing a
  role.
- [Rclean](https://github.com/ProvTools/Rclean) was created to help
  scientists more *easily* write "cleaner" code by providing a simple,
  rigorous way to isolate the minimal code you need in order to
  produce a specific result (e.g. object, plot, output written to
  disk).
- [Rclean](https://github.com/ProvTools/Rclean) is built on data
  provenance, a formal representation of a computation, and uses graph
  analysis to determine the minimal path from inputs to results.
- The goal is to help facilitate code transparency, reproduciblity,
  reusability and ultimately help scientists spend more time on
  research and less time on software.

Install and Setup
=================

You can install
[Rclean](https://cran.r-project.org/web/packages/Rclean/) from *CRAN*:

```{r}
install.packages("Rclean")
```

You can install the most up to date (beta) version easily with
[devtools](https://github.com/hadley/devtools):

```{r}
install.packages("devtools")
devtools::install_github("ProvTools/Rclean", ref = "dev")
```

Once installed, per usual R practice, just load the *Rclean*:

```{r}
library(Rclean)
```

Usage
=====

*Rclean* usage is simple. Have a script with code you want to clean
saved to disk. Then, just run the `clean` function with the path to
the script as the input:

```{r}
clean("./example/simple_script.R")
```

This returns a list of possible results detected in the script,
including execution lines (not counting lines with no code or that are
commented). We can now pick the result we want to focus on for
cleaning:

```{r}
clean("./example/simple_script.R", "tab.15")
```

This produces the minimal code detected from the script. It also
detects library dependencies and inserts them into the code (`libs =
TRUE`).

It can be handy just to take a look at the isolated code, but you can
save the code for later use or sharing (e.g. creating a reproducible
example for getting help) with the `write.code` function:

```{r}
my.code <- clean("./example/simple_script.R", "tab.15")
write.code(my.code, file = "x.R")
```

If you would like to copy your code to your clipboard, you can do that
by not specifying a file path. You can now paste as needed to create a
simpler script.

```{r}
write.code(my.code)
```

## Retrospective Provenance

So far, we've been using "prosepective" provenance generated from the
static code prior to execution. *Rclean* can also be used with
"retrospective" provenance, which is recorded during execution of a
script. Using it facilitates more accurate code cleaning, We can pass
the provenance to the `clean` function via `options`:

```{r}
options(prov.json = readLines("prov_micro.json"))
```

Now that we have the provenance loaded, we can start
cleaning. [Rclean](https://github.com/ProvTools/Rclean) will give us a
list of possible values we can get code for, notice that the option
*rp* (i.e. "retrospective provenance") has been set to `TRUE`:

```{r}
clean(rp = TRUE)

```

Similar to before, you can then pick and choose from among these
results and get the essential code to produce the output, like so:

```{r}
clean(x, rp = TRUE)

```

Happy cleaning!

*Contributing*: if you would like to contribute, please read [[CONTRIBUTE.md]].