### 2019-10-28 inst/README.rmd
roxygenize

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
original_files <- dir()
```

<!-- # ijtiff  <img src="man/figures/logo.png" height="140" align="right"> -->

<!-- Code status -->
[![Build Status](https://travis-ci.org/ProvTools/Rclean.svg?branch=master)](https://travis-ci.org/ProvTools/Rclean) 
[![Coverage status](https://codecov.io/gh/provtools/Rclean/branch/master/graph/badge.svg)](https://codecov.io/github/provtools/Rclean?branch=master)

<!-- R status -->
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/Rclean)](https://cran.r-project.org/package=Rclean)
![RStudio CRAN downloads](http://cranlogs.r-pkg.org/badges/grand-total/Rclean)
![RStudio CRAN monthly downloads](http://cranlogs.r-pkg.org/badges/Rclean)
[![Rdocumentation](http://www.rdocumentation.org/badges/version/Rclean)](http://www.rdocumentation.org/packages/Rclean)

<!-- Dev status -->
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)

<!-- Package Review -->
[![](https://badges.ropensci.org/327_status.svg)](https://github.com/ropensci/software-review/issues/327)
[![status](http://joss.theoj.org/papers/334d80d5508056dc6e7e17c6fd3ed5a6/status.svg)](http://joss.theoj.org/papers/334d80d5508056dc6e7e17c6fd3ed5a6)

<!-- Archiving -->
[![DOI](https://zenodo.org/badge/102645585.svg)](https://zenodo.org/badge/latestdoi/102645585)

Clean up your code
==================

- [Rclean](https://github.com/ProvTools/Rclean) was created to help
  scientists more *easily* write "cleaner" code. 
- Written with research scientists that are results oriented in mind,
  the package's primary function provides a simple way to isolate the
  minimal code you need to produce a specific result, such as a
  statistical table or a figure.
- By focusing on specific results, large and/or complicated analytical
  scripts can be paired down to the essentials and easily re-factored
  to be more robust and easily shared.

Install
=======

You can install
[Rclean](https://cran.r-project.org/web/packages/Rclean/) from *CRAN*:

```{r eval = FALSE}
install.packages("Rclean")
```

You can install the most up to date version easily with
[devtools](https://github.com/hadley/devtools):

```{r eval = FALSE}
install.packages("devtools")
devtools::install_github("MKLau/Rclean")
```

Once installed, per usual R practice, just load the *Rclean* package with:

```{r eval = TRUE}
library(Rclean)
```

Usage
=====

```{r eval = TRUE, echo = FALSE results = "hide"}
library(CodeDepends)
script <- system.file(
    "example", 
    "simple_script.R", 
    package = "Rclean")

```

*Rclean* usage is simple. Have a script with code you want to clean
saved to disk. Then, just run the `clean` function with the path to
the script as the input. Here, we can use an example script that is
included with the package:

```{r eval = FALSE} 
script <- "example/simple_script.R"

```

Here's a quick look at the code:

```{r eval = TRUE}
readLines(script)

```

You can get a list of the variables found in an object with
`get_vars`.

```{r eval = TRUE}
get_vars(script)

```

Sometimes for more complicated scripts, it can be helpful to see a
 network graph showing the interdependencies of
 variables. `code_graph` will produce a network diagram showing which
 lines of code produce which variables (e.g. 1 -> "out"):

```{r eval = TRUE}
code_graph(script)
```

Now, we can now pick the result we want to focus on for cleaning:

```{r eval = TRUE}
clean(script, "tab.15")
```

We can also select several variables at the same time:

```{r eval = TRUE}
my.vars <- c("tab.12", "tab.15")
clean(script, my.vars)
```

While just taking a look at the simplified code can be very helpful,
you can also save the code for later use or sharing (e.g. creating a
reproducible example for getting help) with `keep`:

```{r eval = FALSE}
my.code <- clean(script, my.vars)
keep(my.code, file = "results_tables.R")
```

If you would like to copy your code to your clipboard, you can do that
by not specifying a file path. You can now paste the simplified as needed.

```{r eval = FALSE}
keep(my.code)
```

Contributing
============

Please note that the 'Rclean' project is released with a
[Contributor Code of Conduct](CODE_OF_CONDUCT.md).
By contributing to this project, you agree to abide by its terms.