### 2020-02-13 joss/paper.rmd
Removes multi-cites and brackets

# Introduction

The growth of programming in the sciences has been explosive in the
last decade. This has facilitated the rapid advancement of science
through the agile development of computational tools. However,
concerns have begun to surface about the reproducibility of scientific
research in general @Baker2016 and the potential issues stemming from
issues with analytical software @Stodden2018. Specifically, there is a
growing recognition across disciplines that simply making data and
software "available" is not enough and that there is a need to improve
the transparency and stability of scientific software @Pasquier2018.

At the core of the growth of scientific computation, the ``R``
statistical programming language has grown exponentially to become one
of the top ten programming languages in use today. At it's root R is a
*statistical* programming language. That is, it was designed for use
in analytical workflows; and the majority of the R community is
focused on producing code for idiosyncratic projects that are
*results* oriented. Also, R's design is intentionally at a level that
abstracts many aspects of programming that would otherwise act as a
barrier to entry for many users. This is good in that there are many
people who use R to their benefit with little to no formal training in
computer science or software engineering, but these same users can
also be frequently frustrated by code that is fragile, buggy and
complicated enough to quickly become obtuse even to the authors. The
stability, reproducibility and re-use of scientific analyses in R
would be improved by refactoring, which is a common practice in
software engineering @Martin2009CleanCraftsmanship. From this
perspective, tools that can lower the time and energy required to
refactor analytical scripts and otherwise help to "clean" code, but
abstracted enough to be easily accessible, could have a significant
impact on scientific reproducibility across all disciplines
@Visser2015.

To provide support for easier refactoring in R, we have created
``Rclean``. The ``Rclean`` package provides tools to automatically
reduce a script to the parts that are specifically relevant to a
research product (e.g.  a scientific report, academic talk, research
article, etc.). Although potentially useful to all R coders, it was
designed to ease refactoring for scientists that use R but do not have
formal training in software engineering.

# Methods

The goal of `Rclean` is to provide a set of tools that help someone
reduce and organize code based on results. More often then not, when
someone is writing an R script, the intent is to produce a set of
results, such as a statistical analysis, figure, table, etc. This set
of results is always a subset of a much larger set of possible ways to
explore a dataset, as there are many statistical approaches and tests,
let alone ways to create visualizations and other representations of
patterns in data. This commonly leads to lengthy, complicated scripts
from which researchers manually subset results, but likely never to be
refactored because of the difficulty in disentangling the code needed
to produce some results and not others. The 'Rclean' package uses an
automated technique based on data provenance to analyze existing
scripts and provide ways to identify and extract code to produce a
desired output.

## Data Provenance

All of these processes rely on the generation of data provenance.  The
term provenance means information about the origins of some
object. Data provenance is a formal representation of the execution of
a computational process (https://www.w3.org/TR/prov-dm/), to
rigorously determine the the unique computational pathway from inputs
to results @Carata2014. To avoid confusion, note that "data" in this
context is used in a broad sense to include all of the information
generated during computation, not just the data that are collected in
a research project that are used as input to an analysis. Having the
formalized, mathematically rigorous representation that data
provenance provides guarantees that the analyses that `Rclean`
conducts are theoretically sound. Most importantly, because the
relationships defined by the provenance can be represented as a graph,
it is possible to apply network search algorithms to determine the
minimum and sufficient code needed to generate the chosen result in
the `clean` function.

There are multiple approaches to collecting data provenance, but
`Rclean` uses "prospective" provenance, which analyzes code and uses
language specific information to predict the relationship among
processes and data objects. `Rclean` relies on a library called
`CodeDepends` to gather the prospective provenance for each
script. For more information on the mechanics of the `CodeDepends`
package, see @R-CodeDepends. To get an idea of what data provenance
is, take a look at the `code_graph` function. The plot that it
generates is a graphical representation of the prospective provenance
generated for `Rclean` \ref{fig:prov-graph}.

```{R load-script, echo = FALSE, results = "hide"}
library(CodeDepends)
library(Rclean)
```

```{R prov-graph, fig.cap = "Network diagram of the prospective data provenance generated for an example script. Arrows indicate which lines of code (numbered) produced which objects (named).", echo = FALSE}

script <- system.file(
    "example", 
    "simple_script.R", 
    package = "Rclean")
code_graph(script)

```

In the future, it would also be useful to extend the existing
framework to support other provenance methods. One such possibility is
*retrospective provenance*, which tracks a computational process as it
is executing. Through this active, concurrent monitoring,
retrospective provenance can gather information that static
prospective provenance can't. Greater details of the computational
process would enable other features that could address some
challenges, such as processing information from comments, parsing
control statements and replicating random processes. However, using
retrospective provenance comes at a cost. In order to gather it, the
script needs to be executed. When scripts are computationally
intensive or contain bugs that stop execution, then retrospective
provenance can not be obtained for part or all of the code. Some work
has already been done in the direction of implementing retrospective
provenance for code cleaning in R (see
http://end-to-end-provenance.github.io).

## Software Availability

The software is currently hosted on Github, and we recommend using the
`devtools` library @R-devtools to install directly from the
repository (https://github.com/ROpenSci/Rclean). The package is
open-source and welcomes contributions. Please visit the repository
page to report issues, request features or provide other feedback.

# Discussion

We see promise in connecting `Rclean` with other clean code and
reproducibility tools. One example is the `reprex` package, which
provides a simple API for sharing reproducible examples
@R-reprex. Another possibility is to help transition scripts to
function, package and workflow creation and refactoring via toolboxes
like `drake` @R-drake. `Rclean` could provide a reliable way to
extract parts of a larger script that would be piped to a simplified
reproducible example, in the case of `reprex`, or, since it can
isolate the code from inputs to one or more outputs, be used to
extract all of the components needed to write one or more functions
that would be a part of a package or workflow, as is the goal of
`drake`. To conclude, we hope that `Rclean` makes writing scientific
software easier for the R community. We look forward to feedback and
help with extending its application, particularly in the area of
reproducibility. To get involved, report bugs, suggest features,
please visit the project page.

# Acknowledgments

This work was improved by discussions with ecologists at Harvard
Forest and through the helpful review provided by the ROpenSci
community, particularly Anna Krystalli, Will Landau and Clemens
Schmid. Much of the work was funded by US National Science Foundation
grant SSI-1450277 for applications of End-to-End Data Provenance.

# References

<!-- Use overleaf papers.bib + knitr::write_bib -->