### 2020-03-12 ropensci/rclean_technote.Rmd
Increases figure resolution

---
title: Automatic Code Cleaning in R with `Rclean`
author: M.K. Lau
bibliography: refs.bib
---

> Leave the code cleaner than you found it.

-- R.C. Martin in *Clean Code*

Data analysis can be messy and complicated. Because of project
complexities and time contraints, analytial software often reflects
these challenges. "What did I measure? What analyses are relevant to
them? Do I need to transform the data? What's the function for the
analysis I want to run?" In addition, although many researchers see
the value in learning to write software, the entry level is still
exceedingly high. The `R` language has become very popular among
scientists and analysts because of the lower threshold, which is good
in that it empowers science. The downside is that learning to create
good software is a skill that takes a considerable amount of time to
learn, which means that many scientists write code that will get the
job "done" but lacks transparency and is highly unstable (i.e. buggy).

The goal of `Rclean` is to provide a tool that help someone reduce and
organize code based on results.  More often then not, when someone is
writing an R script, the intent is to produce a set of results, such
as a statistical analysis, figure or table. This set of results is
nearly always a subset of a much larger set of possible ways to
explore a dataset, as there are many statistical approaches and tests,
let alone ways to create visualizations and other representations of
patterns in data. This commonly leads to lengthy, complicated scripts
from which researchers manually subset results, but likely never to be
refactored because of the difficulty in disentangling the code needed
to produce some results and not others.

The `Rclean` package uses an automated technique based on data
provenance and network algorithms to analyze existing scripts and
provide ways to identify and extract code to produce a desired output
without having to manually conduct the search through the code line by
line. `Rclean` was designed to ease refactoring for scientists that
use R but do not have formal training in software engineering and
specifically with the "art" of creating clean code; however, because
manually culling code is tedious and potentially leads to errors,
`Rclean` can be a useful tool for everyone at all levels of coding
ability.

# The Provenance Engine Under the Hood

The process that `Rclean` uses relies on the generation of data
provenance.  The term provenance means information about the origins
of some object. Data provenance is a formal representation of the
execution of a computational process (https://www.w3.org/TR/prov-dm/),
to rigorously determine the the unique computational pathway from
inputs to results [@Carata2014]. To avoid confusion, note that "data"
in this context is used in a broad sense to include all of the
information generated during computation, not just the data that are
collected in a research project that are used as input to an
analysis. Having the formalized, mathematically rigorous
representation that data provenance provides guarantees that the
analyses that `Rclean` conducts are theoretically sound. Most
importantly, because the relationships defined by the provenance can
be represented as a graph, it is possible to apply network search
algorithms to determine the minimum and sufficient code needed to
generate the chosen result in the `clean` function.

There are multiple approaches to collecting data provenance, but
`Rclean` uses "prospective" provenance, which analyzes code and uses
language specific information to predict the relationship among
processes and data objects. `Rclean` relies on a library called
`CodeDepends` to gather the prospective provenance for each
script. For more information on the mechanics of the `CodeDepends`
package, see [@R-CodeDepends]. To get an idea of what data provenance
is, take a look at the `code_graph` function. The plot that it
generates is a graphical representation of the prospective provenance
generated for `Rclean` \ref{fig:prov-graph}.

```{R load-script, echo = FALSE, results = "hide"}
library(CodeDepends)
library(Rclean)
```

```{R prov-graph, dpi = 300, fig.cap = "Network diagram of the prospective data provenance generated for an example script. Arrows indicate which lines of code (numbered) produced which objects (named).", echo = FALSE}

script <- system.file(
    "example", 
    "simple_script.R", 
    package = "Rclean")
code_graph(script)

```

The provenance provides a network representation of relationships
among functions and objects. This is very powerful because we can now
apply algorithms analyze the R script with respect to our results. The
main function of `Rclean`, called `clean`, takes the provenance and
applies a depth first network search algorithm to determine the
pathways leading from inputs to outputs, excluding any objects or
lines of code that do not fall along that pathway and are, therefore,
not necessary to produce the desired set of results. As demonstrated
in the next section, through this process, we can produce a new script
with the minimal code necessary to generate the output we want.

# Installation

Through the helpful feedback from the ROpenSci community, the package
has recently passed software review and a supporting article was
recently published in the Journal of Open Source Software [@Lau2020],
in which you can find more details about the package. The package is
hosted through the ROpenSci organization on Github, and the package
can be installed using the `devtools` library [@R-devtools] directly
from the repository (https://github.com/ROpenSci/Rclean).

```{R install}
library(devtools)
install_github("ROpenSci/Rclean")

```

# Using `Rclean` to clean your code

Analytical scripts that have not been refactored are often both long
and complicated. However, a script doesn't need to be long to be
complicated. The following example script presents some challenges
such that even though it's not a long script, picking through it to
get a result would likely prove to be frustrating.

```{R long-setup, echo = FALSE, eval = TRUE}

script.long <- system.file(
    "example", 
    "long_script.R", 
    package = "Rclean")
script <- script.long

```

```{R long-script, echo = TRUE, eval = FALSE}

library(stats)
x <- 1:100
x <- log(x)
x <- x * 2
x <- lapply(x, rep, times = 4)
### This is a note that I made for myself.
### Next time, make sure to use a different analysis.
### Also, check with someone about how to run some other analysis.
x <- do.call(cbind, x)

### Now I'm going to create a different variable.
### This is the best variable the world has ever seen.

x2 <- sample(10:1000, 100)
x2 <- lapply(x2, rnorm)

### Wait, now I had another thought about x that I want to work through.

x <- x * 2
colnames(x) <- paste0("X", seq_len(ncol(x)))
rownames(x) <- LETTERS[seq_len(nrow(x))]
x <- t(x)
x[, "A"] <- sqrt(x[, "A"])

for (i in seq_along(colnames(x))) {
    set.seed(17)
    x[, i] <- x[, i] + runif(length(x[, i]), -1, 1)
}

### Ok. Now I can get back to x2.
### Now I just need to check out a bunch of stuff with it.

lapply(x2, length)[1]
max(unlist(lapply(x2, length)))
range(unlist(lapply(x2, length)))
head(x2[[1]])
tail(x2[[1]])

## Now, based on that stuff, I need to subset x2.

x2 <- lapply(x2, function(x) x[1:10])

## And turn it into a matrix.
x2 <- do.call(rbind, x2)

## Now, based on x2, I need to create x3.
x3 <- x2[, 1:2]
x3 <- apply(x3, 2, round, digits = 3)

## Oh wait! Another thought about x.

x[, 1] <- x[, 1] * 2 + 10
x[, 2] <- x[, 1] + x[, 2]
x[, "A"] <- x[, "A"] * 2

## Now, I want to run an analysis on two variables in x2 and x3.

fit.23 <- lm(x2 ~ x3, data = data.frame(x2[, 1], x3[, 1]))
summary(fit.23)

## And while I'm at it, I should do an analysis on x.

x <- data.frame(x)
fit.xx <- lm(A~B, data = x)
summary(fit.xx)
shapiro.test(residuals(fit.xx))

## Ah, it looks like I should probably transform A.
## Let's try that.
fit_sqrt_A <- lm(I(sqrt(A))~B, data = x)
summary(fit_sqrt_A)
shapiro.test(residuals(fit_sqrt_A))

## Looks good!

## After that. I came back and ran another analysis with 
## x2 and a new variable.

z <- c(rep("A", nrow(x2) / 2), rep("B", nrow(x2) / 2))
fit_anova <- aov(x2 ~ z, data = data.frame(x2 = x2[, 1], z))
summary(fit_anova)

```

So, let's say we've come to our script wanting to extract the code to
produce one of the results `fit_sqrt_A`, which is an analysis that is
the fitted model object for an analysis. We might want to double check
the results, and we also might need to use the code again for another
purpose, such as creating a plot of the patterns supported by the
test. Manually tracing through our code for all the variables used in
the test and finding all of the lines that were used to prepare them
for the analysis would be annoying and difficult, especially given the
fact that we have used "x" as a prefix for multiple unrelated objects
in the script. However, `Rclean` can do this easily via the `clean`.

```{R long}
library(Rclean)
clean(script, "fit_sqrt_A")

```

As you can see, `Rclean` has picked through the tangled bits of code
and found the minimal set of lines relevant to our object of
interest. This code can now be visually inspected to adapt the
original code or ported to a new, "refactored" script using `keep`.

```{R ex-keep, eval = FALSE}
fitSA <- clean(script, "fit_sqrt_A")
keep(fitSA)

```

This will pass the code to the clipboard for pasting into another
document. To write directly to a new file, a file path can be
specified.

```{R ex-keep-write, eval = FALSE}
fitSA <- clean(script, "fit_sqrt_A")
keep(fitSA, file = "fit_SA.R")

```

To explore more possible variables to extract, the `get_vars` function
can be used to produce a list of the variables (aka. objects) that are
created in the script.

```{R ex-get_vars-hid, eval = TRUE, echo = TRUE}

get_vars(script)

```

Especially when the code for different variables are entangled, it can
be useful to visual the code in order to devise an approach to
cleaning. As seen above \ref{fig:prov-graph}, `code_graph` will
produce a visual of the code and the objects that they produce.

```{R ex-code_graph, dpi = 300, fig.cap = "Example of the plot produced by the code_graph function showing which lines of code produce which variables and which variables are used by other lines of code."}

code_graph(script)

```

After examining the output from `get_vars` and `code_graph`, it is
possible that more than one object needs to be isolated. To do this is
simple in `keep` by passing the set of desired objects to the *vars*
argument.

```{R clean-mlti}

clean(script, vars = c("fit_sqrt_A", "fit_anova"))

```

Currently, libraries can not be isolated directly during the cleaning
process. So, the `get_libs` function provides a way to detect the
libraries for a given script. We just need to supply a file path and
`get_libs` will return the libraries that are called by that script.

```{R ex-get_libs}

get_libs(script)

```

# Concluding Remarks and Future Work

We have created ``Rclean`` to provide a simple, easy to use tool for
scientists who would like help refactoring code. Using ``Rclean`` the
code necessary to produce a specified result (e.g., an object stored
in memory or a table or figure written to disk) can be easily and
*reliably* isolated even when tangled with code for other
results. Tools, such as this, that make it easier to produce
transparent, accessible code will be an important aid for improving
scientific reproducibility.

The `clean` function provides an effective way to remove code that is
unwanted; however, many researchers are wary about doing this exact
thing for at least a few reasons. Perhaps the top reason is that the
main goal of an analysis is the results and taking time to craft
transparent, dependable software is not the priority. As such, taking
time to go back through a script and remove code is time
wasted. Relatedly, for most researchers the best way to keep track of
the various analyses that they have explored is to keep them in the
script, as they do not use a rigorous version control system but
instead rely on file backups and informal versioning. Although we
can't give researchers more hours in the day, providing an easier and
more reliable means to remove unused code will lower the barrier to
creating better, cleaner code. Combined with the increasing use of
version control systems and digital notebooks, the practice of
"saving" analytical ideas in a script will become less common and code
quality will increase.

Relatedly, it is important to point out that `Rclean` *does not* keep
comments present in code. This is a limitation of the data provenance,
which currently does not assign them a relationships. Therefore,
although there is often very useful or even invaluable information in
comments, the `clean` function removes them. This is a general issue
with automated methods for detecting the relationships between
comments and the code itself. Comments at the end of lines are
typically relevant to the line they are on, but this is not a
requirement and could refer to other lines. Also, comments occupying
their own lines usually refer to the following lines, but this is also
not necessarily the case. As `clean` depends on the unambiguous
determination of relationships in the production of results, it cannot
operate automatically on comments. However, comments in the original
code remain untouched and can be used to inform the reduced
code. Also, as the `clean` function is oriented toward isolating code
based on a specific result, the resulting code tends to naturally
support the generation of new comments that are higher level
(e.g. "The following produces a plot of the mean response of each
treatment group."), and lower level comments are not necessary because
the code is simpler and clearer. Lastly, as stated above, comments can
serve an important role in coding, it is worth reflecting on the
statement in R.C. Martin's book `Clean Code: A Handbook of Agile
Software Craftsmanship` that "Comments do not compensate for bad
code."

In the future, it would also be useful to extend the existing
framework to support other provenance methods. One possibility is
*retrospective provenance*, which tracks a computational process as it
is executing. Through this active, concurrent monitoring,
retrospective provenance can gather information that static
prospective provenance can't. Greater details of the computational
process would enable other features that could address some
challenges, such as libraries that are actually used by the code,
processing comments, parsing control statements and replicating random
processes. Using retrospective provenance comes at a cost. In order to
gather it, the script needs to be executed. When scripts are
computationally intensive or contain bugs that stop execution, then
retrospective provenance can not be obtained for part or all of the
code. Some work has already been done in the direction of implementing
retrospective provenance for code cleaning in R (see
http://end-to-end-provenance.github.io). Although such costs may
present challenges, combining prospective and retrospective provenance
methods could provide a powerful and flexible solution.

Although the current implementation of Rclean for minimizing code is
useful on its own, we see promise in connecting it with other clean
code and reproducibility tools. One example is the `reprex` package,
which provides a simple API for sharing reproducible examples
[@R-reprex]. Another possibility is to help transition scripts to
function, package and workflow creation and refactoring via toolboxes
like `drake` [@R-drake]. `Rclean` could provide a reliable way to
extract parts of a larger script that would be piped to a simplified
reproducible example, in the case of `reprex`, or, since it can
isolate the code from inputs to one or more outputs, be used to
extract all of the components needed to write one or more functions
that would be a part of a package or workflow, as is the goal of
`drake`.

To conclude, we hope that `Rclean` makes writing scientific software
easier for the R community. As an open-source project, we look forward
to feedback and help with extending it, particularly in the area of
reproducibility. We invite people to use the package and get involved
by reporting bugs and suggesting or contributing features. For more
information please visit the project page on Github.

# References

<!-- Use overleaf papers.bib + knitr::write_bib -->