### 2020-03-13 ropensci/rclean_technote.Rmd
Edits throughout for final version

---
title: Automatic Code Cleaning in R with `Rclean`
author: M.K. Lau
bibliography: refs.bib
---

# Introduction

> Leave the code cleaner than you found it.

-- R.C. Martin in *Clean Code*

The `R` language has become very popular among scientists and analysts
because it enables the rapid development of software and empowers
scientific investigation. However, regardless of the language used,
data analysis is usually complicated. Because of various project
complexities and time contraints, analytical software often reflects
these challenges. "What did I measure? What analyses are relevant to
them? Do I need to transform the data? What's the function for the
analysis I want to run?" Although many researchers see the value in
learning to write software, the time investment for learning a
programming language alone is still exceedingly high for many, let
alone also learning software best practices. The downside to the rapid
spread of data science is that learning to create good software takes
a back-seat to just writing code that will get the job "done" leading
to issues with transparency and software that is highly unstable
(i.e. buggy and not reproducible).

The goal of `Rclean` is to provide a automated tool to help scientists
more easily write better code. Specifically, `Rclean` has been
designed to faciliate the isolation of the code needed to produce one
or more results, because more often then not, when someone is writing
an R script, the ultimate goal is analytical results for inference,
such as a set of statistical analyses and associated figures and
tables. As the investigative process is inherently iterative, this set
of results is nearly always a subset of a much larger set of possible
ways to explore a dataset. There are many statistical tests and
visualizations and other representations that can be employed in a
myriad of ways depending on how the data are processed. This commonly
leads to lengthy, complicated scripts from which researchers manually
subset results, but likely never to be refactored because of the
difficulty in disentangling the code.

The `Rclean` package uses a technique based on data provenance and
network algorithms to isolate code for a desired result
*automatically*. The intent is to ease refactoring for scientists that
use R but do not have formal training in software design and
specifically with the "art" of creating clean code, which in part is
the development of an intuitive sense of how code is and/or should be
organized. However, manually culling code is tedious and potentially
leads to errors, and we see `Rclean` as a useful tool for everyone at
all levels of coding ability in `R`. Here, we'll cover details on the
implementation and design of the package, a general example of how it
can be used and thoughts on its future development.

# How to Use `Rclean` to Write Cleaner Code

## Installation

Through the helpful feedback from the ROpenSci community, the package
has recently passed software review and a supporting article was
recently published in the Journal of Open Source Software [@Lau2020],
in which you can find more details about the package. The package is
hosted through the ROpenSci organization on Github, and the package
can be installed using the `devtools` library [@R-devtools] directly
from the repository (https://github.com/ROpenSci/Rclean).

```{R install}
library(devtools)
install_github("ROpenSci/Rclean")

```

## Isolating Code for a Set of Results

Analytical scripts that have not been refactored are often both long
and complicated. However, a script doesn't need to be long to be
complicated. The following example script presents some challenges
such that even though it's not a long script, picking through it to
get a result would likely prove to be frustrating.

```{R long-setup, echo = FALSE, eval = TRUE}

script.long <- system.file(
    "example", 
    "long_script.R", 
    package = "Rclean")
script <- script.long

```

```{R long-script, echo = TRUE, eval = FALSE}

library(stats)
x <- 1:100
x <- log(x)
x <- x * 2
x <- lapply(x, rep, times = 4)
### This is a note that I made for myself.
### Next time, make sure to use a different analysis.
### Also, check with someone about how to run some other analysis.
x <- do.call(cbind, x)

### Now I'm going to create a different variable.
### This is the best variable the world has ever seen.

x2 <- sample(10:1000, 100)
x2 <- lapply(x2, rnorm)

### Wait, now I had another thought about x that I want to work through.

x <- x * 2
colnames(x) <- paste0("X", seq_len(ncol(x)))
rownames(x) <- LETTERS[seq_len(nrow(x))]
x <- t(x)
x[, "A"] <- sqrt(x[, "A"])

for (i in seq_along(colnames(x))) {
    set.seed(17)
    x[, i] <- x[, i] + runif(length(x[, i]), -1, 1)
}

### Ok. Now I can get back to x2.
### Now I just need to check out a bunch of stuff with it.

lapply(x2, length)[1]
max(unlist(lapply(x2, length)))
range(unlist(lapply(x2, length)))
head(x2[[1]])
tail(x2[[1]])

## Now, based on that stuff, I need to subset x2.

x2 <- lapply(x2, function(x) x[1:10])

## And turn it into a matrix.
x2 <- do.call(rbind, x2)

## Now, based on x2, I need to create x3.
x3 <- x2[, 1:2]
x3 <- apply(x3, 2, round, digits = 3)

## Oh wait! Another thought about x.

x[, 1] <- x[, 1] * 2 + 10
x[, 2] <- x[, 1] + x[, 2]
x[, "A"] <- x[, "A"] * 2

## Now, I want to run an analysis on two variables in x2 and x3.

fit.23 <- lm(x2 ~ x3, data = data.frame(x2[, 1], x3[, 1]))
summary(fit.23)

## And while I'm at it, I should do an analysis on x.

x <- data.frame(x)
fit.xx <- lm(A~B, data = x)
summary(fit.xx)
shapiro.test(residuals(fit.xx))

## Ah, it looks like I should probably transform A.
## Let's try that.
fit_sqrt_A <- lm(I(sqrt(A))~B, data = x)
summary(fit_sqrt_A)
shapiro.test(residuals(fit_sqrt_A))

## Looks good!

## After that. I came back and ran another analysis with 
## x2 and a new variable.

z <- c(rep("A", nrow(x2) / 2), rep("B", nrow(x2) / 2))
fit_anova <- aov(x2 ~ z, data = data.frame(x2 = x2[, 1], z))
summary(fit_anova)

```

So, let's say we've come to our script wanting to extract the code to
produce one of the results `fit_sqrt_A`, which is an analysis that is
the fitted model object for an analysis. We might want to double check
the results, and we also might need to use the code again for another
purpose, such as creating a plot of the patterns supported by the
test. Manually tracing through our code for all the variables used in
the test and finding all of the lines that were used to prepare them
for the analysis would be annoying and difficult, especially given the
fact that we have used "x" as a prefix for multiple unrelated objects
in the script. However, `Rclean` can do this easily via the `clean`
function.

```{R long}
library(Rclean)
clean(script, "fit_sqrt_A")

```

As you can see, `Rclean` has picked through the tangled bits of code
and found the minimal set of lines relevant to our object of
interest. This code can now be visually inspected to adapt the
original code or ported to a new, "refactored" script using `keep`.

```{R ex-keep, eval = FALSE}
fitSA <- clean(script, "fit_sqrt_A")
keep(fitSA)

```

This will pass the code to the clipboard for pasting into another
document. To write directly to a new file, a file path can be
specified.

```{R ex-keep-write, eval = FALSE}
fitSA <- clean(script, "fit_sqrt_A")
keep(fitSA, file = "fit_SA.R")

```

To explore more possible variables to extract, the `get_vars` function
can be used to produce a list of the variables (aka. objects) that are
created in the script.

```{R ex-get_vars-hid, eval = TRUE, echo = TRUE}

get_vars(script)

```

Especially when the code for different variables are entangled, it can
be useful to visual the code in order to devise an approach to
cleaning. The `code_graph` function can also give us a visual of the
code and the objects that they produce.

```{R ex-code_graph, dpi = 300, fig.cap = "Example of the plot produced by the code_graph function showing which lines of code produce which variables and which variables are used by other lines of code."}

code_graph(script)

```

After examining the output from `get_vars` and `code_graph`, it is
possible that more than one object needs to be isolated. To do this is
simple in `keep` by passing the set of desired objects to the *vars*
argument.

```{R clean-mlti}

clean(script, vars = c("fit_sqrt_A", "fit_anova"))

```

Currently, libraries can not be isolated directly during the cleaning
process. So, the `get_libs` function provides a way to detect the
libraries for a given script. We just need to supply a file path and
`get_libs` will return the libraries that are called by that script.

```{R ex-get_libs}

get_libs(script)

```

# The Provenance Engine Under the Hood

The `clean` function provides an effective way to remove code that is
unwanted; however, many researchers are wary about doing this exact
thing for at least a few reasons. Perhaps the top reason is that the
main goal of an analysis is the results and taking time to craft
transparent, dependable software is not the priority. As such, taking
time to go back through a script and remove code is time
wasted. Relatedly, for most researchers the best way to keep track of
the various analyses that they have explored is to keep them in the
script, as they do not use a rigorous version control system but
instead rely on file backups and informal versioning. Although we
can't give researchers more hours in the day, providing an easier and
more reliable means to remove unused code will lower the barrier to
creating better, cleaner code. Combined with the increasing use of
version control systems and digital notebooks, the practice of
"saving" analytical ideas in a script will become less common and code
quality will increase.

The process that `Rclean` uses relies on the generation of data
provenance.  The term provenance means information about the origins
of some object. Data provenance is a formal representation of the
execution of a computational process (https://www.w3.org/TR/prov-dm/),
to rigorously determine the the unique computational pathway from
inputs to results [@Carata2014]. To avoid confusion, note that "data"
in this context is used in a broad sense to include all of the
information generated during computation, not just the data that are
collected in a research project that are used as input to an
analysis. Having the formalized, mathematically rigorous
representation that data provenance provides guarantees that the
analyses that `Rclean` conducts are theoretically sound. Most
importantly, because the relationships defined by the provenance can
be represented as a graph, it is possible to apply network search
algorithms to determine the minimum and sufficient code needed to
generate the chosen result in the `clean` function.

There are multiple approaches to collecting data provenance, but
`Rclean` uses "prospective" provenance, which analyzes code and uses
language specific information to predict the relationship among
processes and data objects. `Rclean` relies on a library called
`CodeDepends` to gather the prospective provenance for each
script. For more information on the mechanics of the `CodeDepends`
package, see [@R-CodeDepends]. To get an idea of what data provenance
is, we can use `code_graph` function to get a graphical representation
of the prospective provenance generated for `Rclean`
\ref{fig:prov-graph}.

```{R load-script, echo = FALSE, results = "hide"}
library(CodeDepends)
library(Rclean)
```

```{R prov-graph, dpi = 300, fig.cap = "Network diagram of the prospective data provenance generated for an example script. Arrows indicate which lines of code (numbered) produced which objects (named).", echo = FALSE}

script <- system.file(
    "example", 
    "simple_script.R", 
    package = "Rclean")
code_graph(script)

```

All of this work with the provenance is to get the network
representation of relationships among functions and objects. The
provenance network is very powerful because we can now apply
algorithms analyze the R script with respect to our results. This is
what empowers the `clean` function, which takes the provenance and
applies a a network search algorithm to determine the pathways leading
from inputs to outputs. In the process any objects or lines of code
that do not fall along that pathway are by definition not necessary to
produce the desired set of results and can therefore be removed. As
demonstrated in the example, this property of the provenance network
is what facilitates the robust isolation of the minimal code necessary
to generate the output we want.

One important topic to discuss is that `Rclean` *does not* keep
comments present in code. This is the result of a limitation of the
data provenance collection, which currently does not assign them a
relationship in the provenance network. This is a general issue with
detecting the relationships between comments and code. For example,
comments at the end of lines are typically relevant to the line they
are on, but this is not a requirement and could refer to other
lines. Also, comments occupying their own lines usually refer to the
following lines, but this is also not necessarily the case either, as
there is no "linguistic" requirement for this. Although `Rclean`
cannot operate automatically on comments; however, comments in the
original code remain untouched and can be used to inform the reduced
code. Also, as the `clean` function is oriented toward isolating code
based on a specific result, the resulting code tends to naturally
support the generation of new comments that are higher level
(e.g. "The following produces a plot of the mean response of each
treatment group."), and lower level comments are not necessary because
the code is simpler and clearer. This process of comments is an
important part of writing better code. Lastly, although comments can
serve an important role in coding, it is worth reflecting on the
statement in R.C. Martin's book `Clean Code: A Handbook of Agile
Software Craftsmanship` where he writes that, "Comments do not
compensate for bad code."

# Concluding Remarks and Future Work

``Rclean`` provides a simple, easy to use tool for scientists who
would like help refactoring code. Using ``Rclean``, the code necessary
to produce a specified result (e.g., an object stored in memory or a
table or figure written to disk) can be easily and *reliably* isolated
even when tangled with code for other results. Tools, such as this,
that make it easier to produce transparent, accessible code will be an
important aid for improving scientific reproducibility
[@Pasquier2017].

Although the current implementation of Rclean for minimizing code is
useful on its own, we see promise in connecting it with other
reproducibility tools. One example is the `reprex` package, which
provides a simple API for sharing reproducible examples
[@R-reprex]. Another possibility is to help transition scripts to
function, package and workflow creation and refactoring via toolboxes
like `drake` [@R-drake]. `Rclean` could provide a reliable way to
extract parts of a larger script that would be piped to a simplified
reproducible example, in the case of `reprex`, or, since it can
isolate the code from inputs to one or more outputs, be used to
extract all of the components needed to write one or more functions
that would be a part of a package or workflow, as is the goal of
`drake`.

In the future, it would also be useful to extend the existing
framework to support other provenance methods. One possibility is
*retrospective provenance*, which tracks a computational process as it
is executing. Through this active, concurrent monitoring,
retrospective provenance can gather information that static
prospective provenance can't. Greater details of the computational
process would enable other features that could address some
challenges, such as libraries that are actually used by the code,
processing comments (as discussed above), parsing control statements
and replicating random processes. Using retrospective provenance comes
at a cost, however. In order to gather it, the script needs to be
executed. When scripts are computationally intensive or contain bugs
that stop execution, then retrospective provenance can not be obtained
for part or all of the code. Some work has already been done in the
direction of implementing retrospective provenance for code cleaning
in R (see http://end-to-end-provenance.github.io). Although such costs
may present challenges, combining prospective and retrospective
provenance methods could provide a powerful and flexible solution.

We hope that `Rclean` makes writing scientific software easier for the
R community. We look forward to the future progress of the package and
other "code cleaning" tools. As an open-source project, we are very
receptive to feedback and help with extending it. We invite people to
use the package and get involved by reporting bugs and suggesting or
(hopefully) contributing features. For more information please visit
the project page on Github.

# References

<!-- Use overleaf papers.bib + knitr::write_bib -->